Thesis Title

### 2022-02-14 chapters/accronims.tex
finish sequencers

Global Burden of Disease (GBD)
World Health Organization (WHO)
Global Priority Pathogens (GPP)
antimicrobial resistance (AMR)
T2 Magnetic Resonance (T2MR)
polymerase chain reaction (PCR)
pulsed field gel electrophoresis (PFGE)
Multilocus sequence typing (MLST)
basepairs (bp)
Real-time quantitative PCR (qPCR)
Lateral flow immunoassays (LFIA)
ribossomal RNA (rRNA)

>>>>

Global Burden of Disease (GBD)
World Health Organization (WHO)
Global Priority Pathogens (GPP)
antimicrobial resistance (AMR)
T2 Magnetic Resonance (T2MR)
polymerase chain reaction (PCR)
pulsed field gel electrophoresis (PFGE)
Multilocus sequence typing (MLST)
basepairs (bp)
Real-time quantitative PCR (qPCR)
Lateral flow immunoassays (LFIA)
ribossomal RNA (rRNA)
PubMed Central\textsuperscript{\small\textregistered} (PMC)
Public health surveillance systems (PHSS)
dideoxynucleotide triphosphate (ddNTP) 
deoxynucleotide triphosphate (dNTP)
single molecule real-time sequencing (SMRT)

R-PCR real time

### 2022-02-14 chapters/introduction.tex
shove figure 1 in body
finish sequencers

In addition to the emergence of virulent pathogens, the rise of antimicrobial resistance (AMR) poses a major threat to human health around the world. In 2019 there were an estimated 4·95 million deaths associated with bacterial AMR \citep{murray_global_2022}. In 2017, the WHO released The Global Priority Pathogens (GPP) list \citep{world_health_organization_prioritization_2017} to guide discovery, research and development of new antibiotics for drug-resistant bacterial infections. Besides tuberculosis, the global priority due to being the most common and lethal airborne AMR disease worldwide today, responsible for 250 000 deaths each year, it includes 12 groups of pathogens in three priority categories:

\begin{itemize}
  \item{\textbf{Critical priority}}
  \begin{itemize}
    \item{carbapenem-resistant \textit{Acinetobacter baumannii}}
    \item{carbapenem-resistant \textit{Pseudomonas aeruginosa}}
    \item{carbapenem-resistant, third generation cephalosporin-resistant \textit{Enterobacteriaceae}} 
  \end{itemize}
  \item{\textbf{High priority}}
  \begin{itemize}
    \item{vancomycin-resistant \textit{Enterococcus faecium}}
    \item{vancomycin-resistant, methicillin-resistant \textit{Staphylococcus aureus}}
    \item{clarithromycin-resistant \textit{Helicobacter pylori}}
    \item{fluoroquinolone-resistant \textit{Campylobacter} species}
    \item{fluoroquinolone-resistant \textit{Salmonella} species}
    \item{fluoroquinolone-resistant, third generation cephalosporin-resistant, \textit{Neisseria gonorrhoeae}}
  \end{itemize}
  \item{\textbf{Medium priority}}
    \begin{itemize}
    \item{penicillin-non-susceptible \textit{Streptococcus pneumoniae}}
    \item{ampicillin-resistant \textit{Haemophilus influenzae}}
    \item{fluoroquinolone-resistant \textit{Shigella} species} 
  \end{itemize}
\end{itemize}

@@

Sequencing by hybridisation, originally developed in the 1980s, relies on the binding of one strand of DNA to its complementary strand (hybridisation). By repeated hybridisation and washing cycles, it was possible to build larger contiguous sequence information, based upon overlapping information from the probe hybridisation spot, being sensitive to even single-base mismatches when the hybrid region is short or if specialised mismatch detection proteins are present \citep{slatko_overview_2018, detter_nucleic_2014}. Although widely implemented via DNA chips or microarrays, has largely been displaced by other methods, including sequencing by synthesis \citep{goodwin_coming_2016}.

@@

Many of the currently available sequencing by synthesis methods approaches have been described as cyclic array sequencing platforms, as they involve dispersal of target sequences across the surface of a two-dimensional array, followed by sequencing of those targets \citep{hagemann_overview_2015}.

>>>>

In addition to the emergence of virulent pathogens, the rise of antimicrobial resistance (AMR) poses a major threat to human health around the world. In 2019 there were an estimated 4·95 million deaths associated with bacterial AMR \citep{murray_global_2022}. In 2017, the WHO released The Global Priority Pathogens (GPP) list \citep{world_health_organization_prioritization_2017} to guide discovery, research and development of new antibiotics for drug-resistant bacterial infections (see Figure ~\ref{fig:figure1}). Besides tuberculosis, the global priority due to being the most common and lethal airborne AMR disease worldwide today, responsible for 250 000 deaths each year, it includes 12 groups of pathogens in three priority categories.

\begin{figure*}[h!]
\centering
\includegraphics[width=\textwidth]{figures/introduction/Figure 1.png}
\caption{\textbf{World Health Organisation Global Priority Pathogens list.} This catalogue includes, besides \textit{Mycobacterium tuberculosis} considered the number one global priority, a list of twelve microorganisms grouped under three priority tiers according to their antimicrobial resistance: critical (\textit{Acinetobacter baumannii}, \textit{Pseudomonas aeruginosa} and \textit{Enterobacteriaceae}), high (\textit{Enterococcus faecium}, \textit{Helicobacter pylori}, \textit{Salmonella} species, \textit{Staphylococcus aureus}, \textit{Campylobacter} species and \textit{Neisseria gonorrhoeae}), and medium (\textit{Streptococcus pneumoniae}, \textit{Haemophilus influenzae} and \textit{Shigella} species). The major objective was to encourage the prioritisation of funding and incentives, align research and development priorities of public health relevance, and garner global coordination in the fight against antimicrobial resistant bacteria. Adapted from \cite{world_health_organization_prioritization_2017}.}
\label{fig:figure1}
\end{figure*}

@@

Sequencing by hybridisation, also known as sequencing by ligation, originally developed in the 1980s, relies on the binding of one strand of DNA to its complementary strand (hybridisation). By repeated hybridisation and washing cycles, it was possible to build larger contiguous sequence information, based upon overlapping information from the probe hybridisation spot, being sensitive to even single-base mismatches when the hybrid region is short or if specialised mismatch detection proteins are present \citep{slatko_overview_2018, detter_nucleic_2014}. Although widely implemented via DNA chips or microarrays, has largely been displaced by other methods, including sequencing by synthesis \citep{goodwin_coming_2016}.

@@

Many of the currently available sequencing by synthesis methods approaches have been described as cyclic array sequencing platforms, as they involve dispersal of target sequences across the surface of a two-dimensional array, followed by sequencing of those targets \citep{hagemann_overview_2015}. They can be further classified as either single-nucleotide addition or cyclic reversible termination or as single-nucleotide addition \citep{goodwin_coming_2016}.

The first relies on a single signal to mark the incorporation of a dNTP into an elongating strand, avoiding the use of terminators. As a consequence, each of the four nucleotides must be added iteratively to a sequencing reaction to ensure only one deoxynucleotide triphosphate (dNTP) is responsible for the signal. The Roche 454 Life Sciences pysequencing device \footnote{\url{https://web.archive.org/web/20161226040638/http://454.com/}, snapshot from 26 December 2016}, was the first and most popular instrument implementing this technology, but discontinued since 2013 with support to the platform ceasing since 2016. This system distributes template-bound beads into a PicoTiterPlate along with beads containing an enzyme cocktail. As a dNTP is incorporated into a strand, an enzymatic cascade occurs, resulting in a bio-luminescence signal which is captured by a camera, can be attributed to the incorporation of one or more identical dNTPs at a particular bead \citep{goodwin_coming_2016}. The ThermoFisher Ion Torrent system \footnote{\url{https://www.thermofisher.com/pt/en/home/brands/ion-torrent.html}}, released in 2010 and still available today, replaces the optical sensor, using instead  H+ ions that are released as each dNTP is incorporated in the enzymatic cascade, and the consequential change in pH, to detect a signal \citep{goodwin_coming_2016}. Alongside the 454 pyrosequencing system, this system has difficulty in enumerating long repeats, additionally the the throughput of the method depends on the number of wells per chip, ranging from 10 megabases to 1000 megabases of 100 base reads in length , but with a very short run time (three hours) \citep{hagemann_overview_2015, loman_performance_2012}.

The latter is defined by their use of terminator molecules that are similar to those used in the first generation of sequencing, preventing elongation of the DNA molecule, but unlike the first methods it's reversible. To begin the process, a DNA template is primed by a sequence that is complementary to an adapter region, which will initiate polymerase binding to this double-stranded DNA region. During each cycle, a mixture of all four individually labelled and 3'-blocked dNTPs are added. After the incorporation of a single dNTP to each elongating complementary strand, unbound dNTPs are removed and the surface is imaged to identify which dNTP was incorporated at each cluster by optical capture. The fluorophore and blocking group can then be removed and a new cycle can begin \citep{goodwin_coming_2016}. The Illumina systems, which use this technology, accounts for the largest market share for sequencing instruments compared to other platforms\footnote{\url{https://www.forbes.com/companies/illumina/?sh=774358a91aa6}}, allowing paired-end sequencing and having the highest throughput (from 25 million reads for a MiSeq instrument to  1.2 billion reads for a NextSeq instrument\footnote{\url{https://www.illumina.com/systems/sequencing-platforms.html}}), with read lengths ranging from 45 to 300 bases in length with high accuracy, albeit with long running times (4 to 55 hours), rendering this technology a good choice for many sequencing applications where large read length is not required \citep{loman_performance_2012, gupta_chapter_2014, hagemann_overview_2015}.

\subsubsection{The third generation of DNA sequencing}

Despite their wide adoption, second-generation methods require library preparation and an enrichment or amplification step. These steps are time-consuming, introduce biases related to preferential capture or amplification of certain regions, and produce reads with relatively small size, making transversing repetitive genomic regions impossible if they are larger than the read length \citep{hagemann_overview_2015}. Third generation sequencing technologies, also known as long-read sequencing or single-molecule sequencing, are characterised by the generation of ultra-long-reads, albeit at a much lower throughput than the second generation \citep{hoang_long-reads-based_2022}.They also have the potential to go beyond four-base sequencing to reveal genome-wide patterns of methylation and other chemical modifications that control the biology of bacteria or the virulence of pathogens \citep{korlach_going_2012}. Currently, commercial long-read sequencing is supported by two companies: Pacific Biosciences and Oxford Nanopore Technologies.

The basis of Pacific Biosciences sequencers is known as single molecule real-time sequencing (SMRT), which takes place in single use SMRT Cells. These contain multiple immobilised polymerases which, after binding to an adaptor sequence, begins replication incorporating nucleotides with identifying fluorescent labels. The sequence of fluorescence pulses is recorded into a movie which is then converted into a nucleotide sequence. After the polymerase completes replication of one DNA strand, it continues to sequence the opposite adapter and second strand. As a result, it is possible to generate multiple passes of the same template depending on the lifetime of the polymerase \citep{hoang_long-reads-based_2022, loman_twenty_2015}. This technology has an accuracy comparable with the Illumina systems but requires a higher initial investment cost, are much larger machines in comparison with the benchtop counterparts, and have much lower throughput and longer library preparation protocols \citep{hoang_long-reads-based_2022, wenger_accurate_2019}.

Oxford Nanopore Technologies makes the use of nanopores in small, portable single molecule sequencing devices, capable of generating ultra-long sequences in real-time at relatively low cost. Biological nanopores are embedded in solid state membranes within disposable flow cells which, when a DNA strand passes through the pore driven by a motor protein, each nucleotide causes a change in an ionic current across the membrane, which is later base called \citep{hoang_long-reads-based_2022, loman_twenty_2015}. This process is free from florescence labels and amplification requirements, and after one strand is processed, the pore is available to sequence the next available strand. Sequence quality and length depend on the loaded library but are usually much lower than the alternative counterparts, and its troughput is dependent on the number and lifespan of the nanopore within the flowcell, but still much lower than the alternatives. Despite this, its portability, fast advances, and continued improvement of the flowcells make this a fast adopted technology for long-read sequencing.
